Context-dependent consequences of including lagged effects in demographic models

### 2023-07-21 docs/ipm_comparison_ms.tex
update MS formatting and cross refs

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}

@@

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage[noblocks]{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand\Affilfont{\normalsize}
\usepackage{lineno}\linenumbers
\usepackage{sectsty}\sectionfont{\centering}
\usepackage{sectsty}\subsectionfont{\centering}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Context-dependent consequences of including lagged effects in demographic models},
  pdfkeywords={integral projection models, environmental
stochasticity, lagged effects},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\author[1]{Eric R. Scott}
\author[1,2,3]{Emilio M. Bruna}

\affil[1]{Department of Wildlife Ecology and Conservation, University of
Florida, Gainesville, Florida 32611-0430}
\affil[2]{Center for Latin American Studies, University of Florida,
Gainesville, Florida 32611-5530}
\affil[3]{Biological Dynamics of Forest Fragments Project, INPA-PDBFF,
CP 478, Manaus, Amazonas 69011-970 Brazil}

\date{(draft: 20 July 2023)}
\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, sharp corners, frame hidden, breakable]}{\end{tcolorbox}}\fi

@@

We built three classes of IPMs for comparison which each required
different functional forms of their underlying vital rates models. The
simplest IPM was a general, density-independent, deterministic IPM with
four sub-kernels: growth and survival (\(P\), (\textbf{eq:P?})),
fecundity (\(F\), i.e.~production of new seedlings, (\textbf{eq:F?})),
probability of staying a seedling (always 0), and recruitment (\(R\),
i.e.~seedling survival and establishment, (\textbf{eq:R?}))
((\textbf{eq:mature?}; \textbf{eq:sdlg?}), Figure @ref(fig:lifecycle)).
The probability of staying a seedling, was always equal to zero, since
our definition of seedlings was first year plants only.

\[
n(z^{\prime},t+1) = R(z^\prime)n_s(t) + \int_L^U P(z^\prime,z) n(z,t)\;dz
\]\{\#eq:mature\}

\[
n_s(t+1) = \int_L^U F(z) n(z,t) \; dz
\]\{\#eq:sdlg\}

\[
R(z^\prime) = s_sG_s(z^\prime)
\]\{\#eq:R\}

\[
P(z^\prime, z) = s(z)G(z^\prime, z)
\]\{\#eq:P\}

\[
F(z) = p_f(z)f(z)g
\]\{\#eq:F\}

The number and size of mature plants in the next census is determined by
seedlings entering the mature plant population (i.e.~recruitment ) and
survival and growth (or regression) of mature plants
(\textbf{eq:mature?}). Seedlings survive (\(s_s\)) and grow into mature
plants of a particular size (\(G_s(z^\prime)\)) (\textbf{eq:R?}). Mature
plants survive as a function of size (\(s(z)\)), and grow (or regress)
to a new size as a function of their previous size (\(G(z^\prime, z)\))
(\textbf{eq:P?}). Mature plants flower with a probability that is a
function of size (\(p_f(z)\)) and produce a number of seeds as a
function of size (\(f(z)\)), which germinate and establish as seedlings
with probability \(g\) (\textbf{eq:F?}).

@@

To build the general, density-independent, stochastic, kernel-resampled
IPMs, we included environmental stochasticity in all vital rate models
built using the long term demographic datset by adding a random effect
of year (Figure @ref(fig:lifecycle)). The random effect of year was
included using a factor--smooth interaction which allowed the
relationship between plant size and vital rates to vary in functional
form among transition years. The kernel-resampling approach is to
generate kernels corresponding to each transition year in the
demographic dataset using the random smooths for year, and to iterate
the IPM by drawing from these randomly. This is equivalent to the matrix
selection approach for matrix population models described by Caswell
(2001).

For the third method, we modeled the impacts of drought on vital rates
explicitly and created general, density-independent, stochastic,
parameter-resampled IPMs (\emph{sensu} Metcalf et al. (2015)). We
calculated the standardized precipitation evapotranspiraton index (SPEI)
for our site using a published gridded dataset based on ground
measurements (Xavier et al. 2016) as described in Scott et al. (2022).
For all vital rate models fit using the long term demographic dataset,
we modeled delayed effects of SPEI using distributed lag non-linear
models with a maximum lag of 36 months(Scott et al. 2022) (Figure
@ref(fig:lifecycle)). To iterate these parameter-resampled IPMs, a
random sequence of SPEI values was created by sampling years of the
observed monthly SPEI data. Then, 36 month lags are calculated for each
year starting in February (the month of the demographic census). These
values are then used to predict fitted values from the vital rates
models, generating different kernels at each iteration of the IPM. With
this method, the kernels of successive iterations are not entirely
independent because the SPEI values used in calculating vital rates
include values used in the previous two iterations, but they are
ergodic.

@@

For all vital rates estimated using the long term demographic dataset,
the DLNM model fit the best (dAIC = 0) followed by the model with a
random effect of year, followed by the deterministic model (Table
@ref(tab:aic)).\\

Population growth rates were consistently higher in continuous forest
compared to forest fragments across IPM types (Table @ref(tab:lambdas)).

@@

Figure @ref(fig:pop-states) has some interesting things in it:

\begin{itemize}
\item
  for the deterministic IPM (and the kernel-resampled IPM?) there are
  slightly more of the smallest plants and the largest plants in CF
  compared to FF (i.e.~more medium sized plants in FF).
\item
  For the kernel-resampled IPM (random effect of year), the fluctuations
  are extremely similar between CF and FF
\item
  For the parameter-resampled IPM (DLNM) the size structure of the
  population is a LOT more variable in FF. This makes sense as we know
  lagged effects are more important in fragments.
\item
  Also, the fluctuations in size structure in CF do not match the
  fluctuations in FF as well (can see this by the increased spread of
  points in Figure @ref(fig:pop-states)B)
\item
  Also, in the parameter-resampled IPM (and only in this one), we see a
  shift toward smaller plants in FF compared to CF
\end{itemize}

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1389}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2778}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3333}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1111}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1111}}@{}}
\caption{(\#tab:aic) Comparison of vital rate models used to build IPM.
The `Effect of Environment' column describes how environmental effects
were included in models. Those with `none' were used to build
deterministic IPMs; those with a random effect of year were used to
build stochastic, kernel-resampled IPMs; and those with a distributed
lag non-linear model (DLNM) were used to build stochastic,
parameter-resampled IPMs. `edf' is the estimated degrees of freedom of
the penalized GAM. ΔAIC is calculated within each habitat and vital rate
combination. ΔAIC within 2 indicates models are
equivalent.}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Habitat
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Vital Rate
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Effect of Environment
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
edf
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
ΔAIC
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Habitat
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Vital Rate
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Effect of Environment
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
edf
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
ΔAIC
\end{minipage} \\
\midrule()
\endhead
CF & Survival & Random effect of year & 43.26 & 0 \\
CF & Survival & DLNM & 19.72 & 78.92 \\
CF & Survival & None & 4.976 & 260 \\
CF & Growth & Random effect of year & 78.43 & 0 \\
CF & Growth & DLNM & 23.87 & 158.5 \\
CF & Growth & None & 7.81 & 1896 \\
CF & Flowering & DLNM & 19.59 & 0 \\
CF & Flowering & Random effect of year & 17.19 & 1.627 \\
CF & Flowering & None & 7.468 & 381.9 \\
CF & Seedling survival & None & 1 & 0 \\
CF & Seedling survival & Random effect of year & 1.817 & 1.386 \\
CF & Seedling survival & DLNM & 4.008 & 1.528 \\
CF & Seedling growth & Random effect of year & 9.475 & 0 \\
CF & Seedling growth & DLNM & 8.952 & 2.902 \\
CF & Seedling growth & None & 1 & 172.3 \\
FF & Survival & DLNM & 14.95 & 0 \\
FF & Survival & Random effect of year & 19.21 & 35.68 \\
FF & Survival & None & 4.333 & 51.25 \\
FF & Growth & DLNM & 25.18 & 0 \\
FF & Growth & Random effect of year & 37.84 & 200 \\
FF & Growth & None & 5.599 & 382.8 \\
FF & Flowering & DLNM & 20.61 & 0 \\
FF & Flowering & Random effect of year & 13.81 & 27.4 \\
FF & Flowering & None & 5.007 & 101.7 \\
FF & Seedling survival & DLNM & 5.574 & 0 \\
FF & Seedling survival & Random effect of year & 5.088 & 5.721 \\
FF & Seedling survival & None & 1 & 6.491 \\
FF & Seedling growth & Random effect of year & 6.25 & 0 \\
FF & Seedling growth & DLNM & 8.182 & 2.29 \\
FF & Seedling growth & None & 1 & 5.745 \\
\bottomrule()
\end{longtable}

\begin{table}

\caption{table caption}
\centering
\begin{tabular}[t]{lc}
\toprule
IPM Type & mean time (min.)\\
\midrule
Deterministic & 0.02\\
Stochastic, kernel-resampled & 0.07\\
Stochastic, parameter-resampled & 87.12\\
\bottomrule
\end{tabular}
\end{table}

\newpage
\newpage

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4306}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1389}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3611}}@{}}
\caption{(\#tab:lambdas) Population growth rates for continuous forest
(CF) and forest fragments (FF) under different kinds of IPMs with
bootstrapped, bias-corrected, 95\% confidence intervals.}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
IPM
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Habitat
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(\lambda\)
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
IPM
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Habitat
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(\lambda\)
\end{minipage} \\
\midrule()
\endhead
Deterministic & FF & 0.9778 (0.9736, 0.9823) \\
Deterministic & CF & 0.9897 (0.9877, 0.9920) \\
Stochastic, kernel-resampled & FF & 0.9787 (0.9735, 0.9835) \\
Stochastic, kernel-resampled & CF & 0.9913 (0.9892, 0.9939) \\
dlnm & FF & 0.9595 (0.9459, 0.9689) \\
dlnm & CF & 0.9795 (0.9752, 0.9867) \\
\bottomrule()
\end{longtable}

\caption{Lifecycle diagram of \emph{Heliconia acuminata}. Each
transition is associated with an equation for a vital rate function. The
functions shown on the diagram correspond to those used to construct a
general, density-independent, deterministic IPM. The table below shows
the equivalent equations for stochastic, kernel-resampled IPMs and
stochastic, parameter-resampled IPMs.}

\caption{Relative proportions of plant sizes in the first 250 iterations
of the IPM simulations. Stacked area charts (A) show the relative
size/stage distribution of plants in continuous forest (CF, top row) and
forest fragments (FF, bottom row) in each of the three IPMs (columns).
The proportion of each size class in CF and FF for each iteration is
shown in B with the first 30 iterations removed to not include transient
dynamics. A 1:1 line is plotted in black. Size categories include
seedlings (a discrete category in the IPMs), pre-reproductive 1
(log(size) 0--2.5) that have low average survival (\textless{} 0.9) and
a near 0 probability of flowering, pre-reproductive 2 (log(size)
2.5--4.5) that have a higher average survival probabilty (\textgreater{}
0.8) and a near 0 probability of flowering, reproductive 1 (log(size)
4.5--6) that have a high average survival probability (\textgreater0.95)
and a lower flowering probability (\textless{} 0.25), and reproductive 2
(log(size) 6+) that have a high average survival probability
(\textgreater0.95) and higher flowering proability (\textgreater{}
0.2).}

@@

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
]{article}

@@

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage[noblocks]{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand\Affilfont{\normalsize}
\usepackage{lineno}\linenumbers
\usepackage{sectsty}\sectionfont{\centering}
\usepackage{sectsty}\subsectionfont{\centering}
\usepackage{fontspec}
\usepackage{titlesec}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Context-dependent consequences of including lagged effects in demographic models},
  pdfkeywords={integral projection models, environmental
stochasticity, lagged effects},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\author[1]{Eric R. Scott}
\author{María Uriarte}
\author[1,3,5]{Emilio M. Bruna}

\affil[1]{Department of Wildlife Ecology and Conservation, University of
Florida, Gainesville, Florida 32611-0430 USA}
\affil[2]{Department of Ecology, Evolution and Environmental Biology,
Columbia University 1200 Amsterdam Avenue, New York, New York 10027 USA}
\affil[3]{Center for Latin American Studies, University of Florida,
Gainesville, Florida 32611-5530 USA}
\affil[4]{Biological Dynamics of Forest Fragments Project, INPA-PDBFF,
CP 478, Manaus, Amazonas 69011-970 Brazil}

\date{(draft: 21 July 2023)}
\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners, boxrule=0pt, interior hidden, frame hidden]}{\end{tcolorbox}}\fi

@@

We built three classes of IPMs for comparison which each required
different functional forms of their underlying vital rates models. The
simplest IPM was a general, density-independent, deterministic IPM with
four sub-kernels: growth and survival (\(P\), Equation~\ref{eq-P}),
fecundity (\(F\), i.e.~production of new seedlings,
Equation~\ref{eq-F}), probability of staying a seedling (always 0), and
recruitment (\(R\), i.e.~seedling survival and establishment,
Equation~\ref{eq-R}) (Equation~\ref{eq-mature}, Equation~\ref{eq-sdlg},
Figure~\ref{fig-lifecycle}). The probability of staying a seedling, was
always equal to zero, since our definition of seedlings was first year
plants only.

\begin{equation}\protect\hypertarget{eq-mature}{}{
n(z^{\prime},t+1) = R(z^\prime)n_s(t) + \int_L^U P(z^\prime,z) n(z,t)\;dz
}\label{eq-mature}\end{equation}

\begin{equation}\protect\hypertarget{eq-sdlg}{}{
n_s(t+1) = \int_L^U F(z) n(z,t) \; dz
}\label{eq-sdlg}\end{equation}

\begin{equation}\protect\hypertarget{eq-R}{}{
R(z^\prime) = s_sG_s(z^\prime)
}\label{eq-R}\end{equation}

\begin{equation}\protect\hypertarget{eq-P}{}{
P(z^\prime, z) = s(z)G(z^\prime, z)
}\label{eq-P}\end{equation}

\begin{equation}\protect\hypertarget{eq-F}{}{
F(z) = p_f(z)f(z)g
}\label{eq-F}\end{equation}

The number and size of mature plants in the next census is determined by
seedlings entering the mature plant population (i.e.~recruitment ) and
survival and growth (or regression) of mature plants
(Equation~\ref{eq-mature}). Seedlings survive (\(s_s\)) and grow into
mature plants of a particular size (\(G_s(z^\prime)\))
(Equation~\ref{eq-R}). Mature plants survive as a function of size
(\(s(z)\)), and grow (or regress) to a new size as a function of their
previous size (\(G(z^\prime, z)\)) (Equation~\ref{eq-P}). Mature plants
flower with a probability that is a function of size (\(p_f(z)\)) and
produce a number of seeds as a function of size (\(f(z)\)), which
germinate and establish as seedlings with probability \(g\)
(Equation~\ref{eq-F}).

@@

To build the general, density-independent, stochastic, kernel-resampled
IPMs, we included environmental stochasticity in all vital rate models
built using the long term demographic datset by adding a random effect
of year (Figure~\ref{fig-lifecycle}). The random effect of year was
included using a factor--smooth interaction which allowed the
relationship between plant size and vital rates to vary in functional
form among transition years. The kernel-resampling approach is to
generate kernels corresponding to each transition year in the
demographic dataset using the random smooths for year, and to iterate
the IPM by drawing from these randomly. This is equivalent to the matrix
selection approach for matrix population models described by Caswell
(2001).

For the third method, we modeled the impacts of drought on vital rates
explicitly and created general, density-independent, stochastic,
parameter-resampled IPMs (\emph{sensu} Metcalf et al. (2015)). We
calculated the standardized precipitation evapotranspiraton index (SPEI)
for our site using a published gridded dataset based on ground
measurements (Xavier et al. 2016) as described in Scott et al. (2022).
For all vital rate models fit using the long term demographic dataset,
we modeled delayed effects of SPEI using distributed lag non-linear
models with a maximum lag of 36 months (Scott et al. 2022)
(Figure~\ref{fig-lifecycle}). To iterate these parameter-resampled IPMs,
a random sequence of SPEI values was created by sampling years of the
observed monthly SPEI data. Then, 36 month lags are calculated for each
year starting in February (the month of the demographic census). These
values are then used to predict fitted values from the vital rates
models, generating different kernels at each iteration of the IPM. With
this method, the kernels of successive iterations are not entirely
independent because the SPEI values used in calculating vital rates
include values used in the previous two iterations, but they are
ergodic.

@@

For all vital rates estimated using the long term demographic dataset,
the DLNM model fit the best (dAIC = 0) followed by the model with a
random effect of year, followed by the deterministic model
(Table~\ref{tbl-aic}).\\

Population growth rates were consistently higher in continuous forest
compared to forest fragments across IPM types (Table~\ref{tbl-lambdas}).

@@

Figure~\ref{fig-pop-states} has some interesting things in it:

\begin{itemize}
\item
  for the deterministic IPM (and the kernel-resampled IPM?) there are
  slightly more of the smallest plants and the largest plants in CF
  compared to FF (i.e.~more medium sized plants in FF).
\item
  For the kernel-resampled IPM (random effect of year), the fluctuations
  are extremely similar between CF and FF
\item
  For the parameter-resampled IPM (DLNM) the size structure of the
  population is a LOT more variable in FF. This makes sense as we know
  lagged effects are more important in fragments.
\item
  Also, the fluctuations in size structure in CF do not match the
  fluctuations in FF as well (can see this by the increased spread of
  points in Figure~\ref{fig-pop-states} (B)
\item
  Also, in the parameter-resampled IPM (and only in this one), we see a
  shift toward smaller plants in FF compared to CF
\end{itemize}

@@

@@

\hypertarget{tbl-aic}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1389}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.2778}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3333}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1111}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1111}}@{}}
\caption{\label{tbl-aic}Comparison of vital rate models used to build
IPM. The `Effect of Environment' column describes how environmental
effects were included in models. Those with `none' were used to build
deterministic IPMs; those with a random effect of year were used to
build stochastic, kernel-resampled IPMs; and those with a distributed
lag non-linear model (DLNM) were used to build stochastic,
parameter-resampled IPMs. `edf' is the estimated degrees of freedom of
the penalized GAM. ΔAIC is calculated within each habitat and vital rate
combination. ΔAIC within 2 indicates models are
equivalent.}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Habitat
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Vital Rate
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Effect of Environment
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
edf
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
ΔAIC
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
Habitat
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Vital Rate
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Effect of Environment
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
edf
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
ΔAIC
\end{minipage} \\
\midrule()
\endhead
CF & Survival & Random effect of year & 43.26 & 0 \\
CF & Survival & DLNM & 19.72 & 78.92 \\
CF & Survival & None & 4.976 & 260 \\
CF & Growth & Random effect of year & 78.43 & 0 \\
CF & Growth & DLNM & 23.87 & 158.5 \\
CF & Growth & None & 7.81 & 1896 \\
CF & Flowering & DLNM & 19.59 & 0 \\
CF & Flowering & Random effect of year & 17.19 & 1.627 \\
CF & Flowering & None & 7.468 & 381.9 \\
CF & Seedling survival & None & 1 & 0 \\
CF & Seedling survival & Random effect of year & 1.817 & 1.386 \\
CF & Seedling survival & DLNM & 4.008 & 1.528 \\
CF & Seedling growth & Random effect of year & 9.475 & 0 \\
CF & Seedling growth & DLNM & 8.952 & 2.902 \\
CF & Seedling growth & None & 1 & 172.3 \\
FF & Survival & DLNM & 14.95 & 0 \\
FF & Survival & Random effect of year & 19.21 & 35.68 \\
FF & Survival & None & 4.333 & 51.25 \\
FF & Growth & DLNM & 25.18 & 0 \\
FF & Growth & Random effect of year & 37.84 & 200 \\
FF & Growth & None & 5.599 & 382.8 \\
FF & Flowering & DLNM & 20.61 & 0 \\
FF & Flowering & Random effect of year & 13.81 & 27.4 \\
FF & Flowering & None & 5.007 & 101.7 \\
FF & Seedling survival & DLNM & 5.574 & 0 \\
FF & Seedling survival & Random effect of year & 5.088 & 5.721 \\
FF & Seedling survival & None & 1 & 6.491 \\
FF & Seedling growth & Random effect of year & 6.25 & 0 \\
FF & Seedling growth & DLNM & 8.182 & 2.29 \\
FF & Seedling growth & None & 1 & 5.745 \\
\bottomrule()
\end{longtable}

\hypertarget{tbl-time}{}
\begin{table}
\caption{\label{tbl-time}Figure caption to be written }\tabularnewline

\centering
\begin{tabular}[t]{lc}
\toprule
IPM Type & mean time (min.)\\
\midrule
Deterministic & 0.02\\
Stochastic, kernel-resampled & 0.07\\
Stochastic, parameter-resampled & 87.12\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{tbl-lambdas}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4306}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1389}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3611}}@{}}
\caption{\label{tbl-lambdas}Population growth rates for continuous
forest (CF) and forest fragments (FF) under different kinds of IPMs with
bootstrapped, bias-corrected, 95\% confidence intervals.}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
IPM
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Habitat
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(\lambda\)
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
IPM
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Habitat
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(\lambda\)
\end{minipage} \\
\midrule()
\endhead
Deterministic & FF & 0.9778 (0.9736, 0.9823) \\
Deterministic & CF & 0.9897 (0.9877, 0.9920) \\
Stochastic, kernel-resampled & FF & 0.9787 (0.9735, 0.9835) \\
Stochastic, kernel-resampled & CF & 0.9913 (0.9892, 0.9939) \\
dlnm & FF & 0.9595 (0.9459, 0.9689) \\
dlnm & CF & 0.9795 (0.9752, 0.9867) \\
\bottomrule()
\end{longtable}

\caption{\label{fig-lifecycle}Lifecycle diagram of \emph{Heliconia
acuminata}. Each transition is associated with an equation for a vital
rate function. The functions shown on the diagram correspond to those
used to construct a general, density-independent, deterministic IPM. The
table below shows the equivalent equations for stochastic,
kernel-resampled IPMs and stochastic, parameter-resampled IPMs.}

\caption{\label{fig-pop-states}Relative proportions of plant sizes in
the first 250 iterations of the IPM simulations. Stacked area charts (A)
show the relative size/stage distribution of plants in continuous forest
(CF, top row) and forest fragments (FF, bottom row) in each of the three
IPMs (columns). The proportion of each size class in CF and FF for each
iteration is shown in B with the first 30 iterations removed to not
include transient dynamics. A 1:1 line is plotted in black. Size
categories include seedlings (a discrete category in the IPMs),
pre-reproductive 1 (log(size) 0--2.5) that have low average survival
(\textless{} 0.9) and a near 0 probability of flowering,
pre-reproductive 2 (log(size) 2.5--4.5) that have a higher average
survival probabilty (\textgreater{} 0.8) and a near 0 probability of
flowering, reproductive 1 (log(size) 4.5--6) that have a high average
survival probability (\textgreater0.95) and a lower flowering
probability (\textless{} 0.25), and reproductive 2 (log(size) 6+) that
have a high average survival probability (\textgreater0.95) and higher
flowering proability (\textgreater{} 0.2).}
